#Project 1
------

###by. Kevin Wunderlich

```{r,echo=TRUE}
Proj1 <- read.csv("~/activity.csv")
library(lubridate)
library(lattice)
Proj1$date<-ymd(Proj1$date)

Proj1$WeekDay<-weekdays(x = Proj1$date)

gsub(pattern = "Monday",replacement = "Weekday",x = Proj1$WeekDay)->Proj1$WeekDay
gsub(pattern = "Tuesday",replacement = "Weekday",x = Proj1$WeekDay)->Proj1$WeekDay
gsub(pattern = "Wednesday",replacement = "Weekday",x = Proj1$WeekDay)->Proj1$WeekDay
gsub(pattern = "Thursday",replacement = "Weekday",x = Proj1$WeekDay)->Proj1$WeekDay
gsub(pattern = "Friday",replacement = "Weekday",x = Proj1$WeekDay)->Proj1$WeekDay
gsub(pattern = "Saturday",replacement = "Weekend",x = Proj1$WeekDay)->Proj1$WeekDay
gsub(pattern = "Sunday",replacement = "Weekend",x = Proj1$WeekDay)->Proj1$WeekDay
```

Above I have:
-read in the data
-changed format of the date to year-month-day format using the lubridate package
-by manipulating the weekdays() function, I was able two great two variables "Weekday" and "Weekend" for later reasons.

### What is the mean total and total of steps taken per day?

First, I calculated and plotted the total number of steps for each day

```{r,echo=TRUE}

as.vector(tapply(X = Proj1$steps,INDEX = Proj1$date,FUN = sum))->stepDayTot

hist(stepDayTot,breaks = 15,col = "black",xlab = "Total Steps",main = "Histogram of Total Steps by Day",border = "white")

```

Similarily, I used tapply() function again to calculate the median and mean for each day

####Means by Day
```{r, echo=TRUE}
as.vector(tapply(X = Proj1$steps,INDEX = Proj1$date,FUN = mean))->stepDayAvg
hist(stepDayAvg,breaks = 15,col = "black",xlab = "Day Recorded (days 1-61)",main = "Histogram of Mean Steps by Day",border = "red")
```

####Median by Day

```{r, echo=TRUE}
as.vector(tapply(X = Proj1$steps,INDEX = Proj1$date,FUN = median))->stepDayMed
summary(stepDayMed)
```

Clearly, from the Median read outs, we can see that there are many missing values, which poses many problems, however for the next problem consier NA=0.

###What is the average daily activity pattern

Now it is important to consider the entire data set as many intervals of multiple days, i.e. a time series.

Therefore, we will have to condence the data according to time intervals and may arrive at either plot:


##1) If we consider x-axis to be time intervals 1:288

```{r, echo=TRUE}
Proj1$steps[is.na(Proj1$steps)]<-0
as.vector(tapply(X = Proj1$steps,INDEX = Proj1$interval,FUN = mean))->IntervalSteps
plot(IntervalSteps,type="l",main = "Average Daily Activity",xlab = "Time Intervals",ylab = "Step Count",col="black")
```

or


##2) if we consider x=time intervals 0000:2355 in 5 minute increments

The following plot, allows us to realize that typically around 9am this person is most active.

```{r,echo=TRUE}
plot(x=unique(Proj1$interval),y=IntervalSteps[!is.na(IntervalSteps)],type="l",main = "Average Daily Activity",xlab = "Time Intervals",ylab = "Step Count",col="black")
```

###Now consider all of the missing values:

```{r,echo=TRUE}
sum(is.na(Proj1$steps))->TotalNA
```

There are `r TotalNA ` missing values!!!

My strategy to fix this is to generate random integers with a mean equal to the mean of the daily averages and a standard deviation of 20 steps, and the code follows along with the renewed computations from part 1.

```{r,echo=TRUE}

Proj1$steps[is.na(Proj1$steps)]<-as.integer(rnorm(length(Proj1$steps[is.na(Proj1$steps)]),mean = mean(stepDayAvg,na.rm = TRUE),sd = 20))
Proj1$steps<-abs(Proj1$steps)
as.vector(tapply(X = Proj1$steps,INDEX = Proj1$date,FUN = sum))->stepDayTot1

hist(stepDayTot1,breaks = 15,col = "black",xlab = "Total Steps",main = "Histogram of Total Steps by Day",border = "white")

as.vector(tapply(X = Proj1$steps,INDEX = Proj1$date,FUN = mean))->stepDayAvg1

hist(stepDayAvg1,breaks = 15,col = "black",xlab = "Mean Steps",main = "Histogram of Mean Steps by Day",border = "red")

as.vector(tapply(X = Proj1$steps,INDEX = Proj1$date,FUN = median))->stepDayMed1
summary(stepDayMed1)
```
 
By replacing these values, we may have a better idea of the subjects activities and even when recording began



###Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE}
xyplot(Proj1$steps~Proj1$interval|Proj1$WeekDay,type='l',xlab = "Time Interval",ylab = "Steps Taken",main="Steps Taken Weekend vs Weekday")

```

It is clear that there is more morning activity during the week than during the weekend.





